---
title: "TDP-43 cross-regulation"
author: "Jack Humphrey"
date: "12/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/google_drive/Work/PhD_Year_4/FUS_meta-analysis/")
library(readr)
library(tidyr)
library(purrr)
library(dplyr)
library(stringr)
```

There are clear TDP-43 iCLIP peaks in the FUS retained introns. This suggests that TDP-43 cross-regulates FUS. Are there changes in FUS intron retention in our TDP-43 mouse data?
For each dataset, load in the SGSeq results and find the FUS introns and exon 7 skipping
Extract PSIs for case and control
Make a set of plots using existing code.

```{r}
i6 <- "chr7:127972770-127974400"
i7 <- "chr7:127974434-127975888"
e7 <- "chr7:127972770-127975888"
```



## Cleveland TDP-43 knockdown


```{r}
cleveland <- "data/SGSeq/TDP/sgseq_Cleveland_TDP43_KD.tab"
d <- read_tsv(cleveland)

fus <- filter(d, geneName == "Fus")

# extract intron 6 and 7


intron6 <- filter(fus, coords == i6, variantType == "RI:R")
intron7 <- filter(fus, coords == i7, variantType == "RI:R")
exon7   <- filter(fus, coords == e7, variantType == "SE:I+RI:R+A5SS:P+A3SS:P")

getPSI <- function(df, event){
  psi <- select(df, ends_with("_psi")) %>%
    tidyr::gather(key = "sample", value = "psi") %>%
    mutate(condition = str_split_fixed(sample, pattern = "_", n = 3)[,1] ) %>%
    mutate( event = event, padj = df$padj, pvalue = df$pvalue) %>%
    mutate(condition = case_when(
      condition == "Control" ~ "CTL",
      condition == "MUT" ~ "TDP-43 KD"
    )) 
  
}

cleveland_psi <- 
  list(intron6, intron7, exon7) %>%
  map2_df( .y = c("intron6", "intron7", "exon7"), getPSI)



```

## M323K adult

```{r}
m323k_adult <- "data/SGSeq/TDP/M323K_New_CONTROL_HOM_res_clean_novel.tab"
d <- read_tsv(m323k_adult)
fus <- filter(d, external_gene_ID == "Fus")

intron6 <- filter(fus, coords == i6, variantType == "RI:R")
intron7 <- filter(fus, coords == i7, variantType == "RI:R")
exon7   <- filter(fus, coords == e7, variantType == "SE:I+RI:R+A5SS:P+A3SS:P")

getPSI <- function(df, event){
  psi <- select(df, ends_with("_psi")) %>%
    tidyr::gather(key = "sample", value = "psi") %>%
    mutate(condition = str_split_fixed(sample, pattern = "_", n = 4)[,2] ) %>%
    mutate(condition = case_when(
      condition == "WT" ~ "CTL",
      condition == "HOM" ~ "M323K HOM"
    ))  %>%
    mutate( event = event, padj = df$padj, pvalue = df$pvalue)
  
}

m323k_adult_psi <- 
  list(intron6, intron7, exon7) %>%
  map2_df( .y = c("intron6", "intron7", "exon7"), getPSI)


```

## F210I embryo

```{r}
f210i_embryo <- "data/SGSeq/TDP/sgseq_f210i_embryonic_brain_hom.tab"
d <- read_tsv(f210i_embryo)

fus <- filter(d, external_gene_name == "Fus") %>%
  mutate( coords = paste0(chrom, ":", chromStart, "-", chromEnd) )

intron6 <- filter(fus, coords == i6, variantType == "RI:R")
intron7 <- filter(fus, coords == i7, variantType == "RI:R")
exon7   <- filter(fus, coords == e7, variantType == "SE:I+RI:R+A5SS:P+A3SS:P")

getPSI <- function(df, event){
  psi <- select(df, ends_with("norm_psi")) %>%
    tidyr::gather(key = "sample", value = "psi") %>%
    mutate(condition = str_split_fixed(sample, pattern = "_", n = 4)[,2] ) %>%
    mutate(condition = case_when(
      condition == "CTL" ~ "CTL",
      condition == "HOM" ~ "F210I HOM"
    ))  %>%
    mutate( event = event, padj = df$padj, pvalue = df$pvalue)
}

f210i_embryo_psi <- 
  list(intron6, intron7, exon7) %>%
  map2_df( .y = c("intron6", "intron7", "exon7"), getPSI)



```

## F210I adult

```{r}
f210i_adult <- "data/SGSeq/TDP/F210I_adult_sc_res_clean_novel.tab"
d <- read.table(f210i_adult,header=TRUE)

fus <- filter(d, geneName == "Fus") %>%
  mutate( chrom = str_split_fixed(from, pattern = ":", 4)[,2],
          chromStart = str_split_fixed(from, pattern = ":", 4)[,3],
          chromEnd = str_split_fixed(to, pattern = ":", 4)[,3] ) %>%
  mutate( coords = paste0(chrom, ":", chromStart, "-", chromEnd) )

intron6 <- filter(fus, coords == i6, variantType == "RI:R")
intron7 <- filter(fus, coords == i7, variantType == "RI:R")
exon7   <- filter(fus, coords == e7, variantType == "SE:I")

getPSI <- function(df, event){
  psi <- select(df, ends_with("_psi")) %>%
    tidyr::gather(key = "sample", value = "psi") %>%
    mutate(condition = str_split_fixed(sample, pattern = "_", n = 4)[,2] ) %>%
    mutate( event = event, padj = df$padj)
}

f210i_adult_psi <- 
  getPSI(exon7, "exon7")

 # map2_df( .y = c("exon7"), getPSI)
#

```


Plot PSIs for each dataset together

```{r}
library(ggplot2)
library(ggbeeswarm)
library(patchwork)

psi_plot <- function(res, mytitle, mysub){
  means <- res %>%
    group_by(condition, event) %>%
    summarise(mean = mean(psi))
  
  res <- left_join(res, means, by = c("condition", "event"))
  
  res <- mutate(res, padj = paste0("P=",signif(padj, digits=2)))
  
  res <- mutate(res, pvalue = paste0("P=",signif(pvalue, digits=2)))
  
  ggplot(data = res) + 
  geom_beeswarm( aes(x = condition, y = psi),priority = "ascending",cex = 5) +
  geom_errorbar( aes(x = condition, ymin = mean, ymax = mean ), width = 0.5 ) +
  facet_wrap(~event, nrow = 3) +
  guides(colour = FALSE) +
  xlab("") +
  theme_bw() +
  ylab("PSI") +
  theme(
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, face = "italic"),
   # strip.background = element_rect(colour="black", fill="white"),
    strip.background = element_blank(),
    axis.text.x = element_text(colour = "black")) +
  geom_text(aes(x = 1.5, y = 1.1, label = pvalue)) +
  ylim(0,1.19) +
  labs(title = mytitle,subtitle = mysub)
    #+
  # scale_colour_manual(values = cols) +
  # coord_cartesian(#ylim = c(0, 0.12), # This focuses the x-axis on the range of interest
  #                 clip = 'off' # This keeps the labels from disappearing
  # )# +#
  # geom_text( aes( x = Condition, y = -0.03, label = Treatment ) ) +
  # annotate(geom = "text", x = -0.01, y = -0.03, label = "CHX") +
  # scale_x_discrete(labels = c("CTL\n","CTL\n","MUT\n","MUT\n"))
}

p <- 
  psi_plot(cleveland_psi, "TDP-43 ASO knockdown", "adult") + 
  psi_plot(m323k_adult_psi, "TDP-43 M323K", "adult") + 
  psi_plot(f210i_embryo_psi, "TDP-43 F210I", "embryo")

p

ggsave(p, filename = "figures/Autoregulation/TDP_cross_regulation.pdf")
```


